---
title: "MSNA AoK Aggregation"
author: "HK"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(openxlsx)
library(tidyverse)
library(data.table)
library(fauxnaif)

source("C:\\Users\\hanna.kurovska\\OneDrive - World Food Programme\\Desktop\\RAM\\R\\functions.R")
```

Dataset and questionnaire

```{r}
ds <- read.xlsx("C:\\Users\\hanna.kurovska\\World Food Programme\\Ukraine Response_ Programme Team - RAM\\04 - Needs assessments\\AoK_2023\\05_Data_cleaning\\Dataset\\AoK_23_Dataset_clean_230911.xlsx")

questions <- read.xlsx("C:\\Users\\hanna.kurovska\\OneDrive - World Food Programme\\Desktop\\RAM\\AoK\\2023\\Data_cleaning\\AoK_23_tool_draft_HK_1_clean.xlsx")
```

Spreadsheet with aggregation rules (priorities)

```{r}
aggr_rules <- read.xlsx("C:\\Users\\hanna.kurovska\\OneDrive - World Food Programme\\Desktop\\RAM\\AoK\\2023\\Aggregation\\Aggregation_rules (1).xlsx")
```

Removing "don't know" prior to calculating crosstabs (replacing with NA)

Select one

```{r}
names(aggr_rules) <- make.unique(names(aggr_rules))
dks <- aggr_rules |> 
  filter(ignore_dk == 1) |>
  select(indicator, var.x, option)

dk_so <- dks |>
  filter(var.x != "1") |>
  select(var.x) |>
  unlist() |>
  as.vector()
  
  
ds <- ds |> mutate(across(!starts_with("_"), ~na_if_in(., unique(dk_so))))
```


Select multiple

If "dk" is selected, the whole sm variable is turned to NA

```{r}
dk_sm <- dks |>
  filter(var.x == "1") |>
  mutate(nam = paste0(indicator, "/", option)) |>
  select(nam) |>
  unlist() |>
  as.vector()

dk_sm_a <- dks |>
  filter(var.x == "1") |>
  mutate(nam = paste0(indicator, "/", option))

for(var in dk_sm){
  
  dk_sm_g <- str_split(var, pattern = "\\/")[[1]][1]
  dk_sm_g_vars <- names(ds)[grepl(paste0(dk_sm_g), names(ds))]
  
  for(i in 1:nrow(ds)){
    if(ds[i, paste0(var)] == 1 | is.na(ds[i, paste0(var)])){
       ds[i, dk_sm_g_vars] <- NA
    }
  }
}
```


Getting settlement-level frequency tables for aggregation

1. getting dependent vars for crosstabulation

```{r}
aggr_var <- "a_5_settlement"
tech_vars <- c("start", "end", "date_survey", "device_id", "g_4_enumerator_comment", "interview_code", "instanceID")
vars <- names(ds)
vars <- vars[!str_starts(vars, "_")]
vars <- setdiff(vars, c(aggr_var, tech_vars))
```

Crosstabs

```{r}
tabs <- get_crosstabs(ds = ds, ind_var = aggr_var, dep_vars = vars)
```

Removing overalls from frequency tables. Splitting tabs into SO and SM questions

```{r}
tabs_sm <- tabs |>
  select(indicator, var, starts_with("prop_")) |>
  select(!contains("Overall")) |>
  filter(str_detect(indicator, "\\/"))

names(tabs_sm) <- str_remove(names(tabs_sm), "prop_")


tabs_so <- tabs |>
  select(indicator, var, starts_with("prop_")) |>
  select(!contains("Overall")) |>
  filter(!str_detect(indicator, "\\/"))

names(tabs_so) <- str_remove(names(tabs_so), "prop_")
```



### Multiple choices aggregation

#### Aggregation rules

"Selected" if >= the threshold (40%)

```{r}
threshold <- 0.4
```

```{r}
# getting settlements names
settlements <- names(tabs_sm)[grepl("UA", names(tabs_sm))]
# comparing with threshold and getting the tabs with replaced values
tabs_sm_aggr <- tabs_sm |>
  mutate(across(all_of(settlements), ~ifelse(.>=threshold, 1, 0)))
# NAs in tabs are 0s
tabs_sm_aggr <- replace(tabs_sm_aggr, is.na(tabs_sm_aggr), 0)
# filtering 1s (selected)
tabs_sm_aggr <- tabs_sm_aggr |>
  filter(var == 1) |>
  select(-var)
#transposing
ds_sm_aggr <- transpose(tabs_sm_aggr)
rownames(ds_sm_aggr) <- colnames(tabs_sm_aggr)
ds_sm_aggr <- setNames(ds_sm_aggr, ds_sm_aggr[1,])
ds_sm_aggr <- ds_sm_aggr[-1,]
ds_sm_aggr <- tibble::rownames_to_column(ds_sm_aggr, "settlement")
```

### Select one questions aggregation

#### Aggregation rules:

for ordinal or quasi-ordinal scales: if 50%+ in one of options then this one, if there is no such option - average of priorities/severities of selected options is taken, and the option with this level of priority/severity is taken. 

for nominal scales: prevaling option. If same %, then - "unknown"


0. Joining the doc with scale type and priority lvls

```{r}
tabs_so_p <- tabs_so |> left_join(aggr_rules |> select(indicator, option, var.x, `Priority.(opposite.of.severity)`, scale), by = c("indicator", "var" = "var.x"))
```


1. Ordinal scales
```{r}
tabs_so_p_ord <- tabs_so_p |> filter(scale == "ordinal")

ds_ord_aggr <- NULL

for(sttl in settlements){

  
  temp_settl <- tabs_so_p_ord |>
  select(indicator, var, !!sym(sttl), `Priority.(opposite.of.severity)`)
# case 1 - 50+%
case1 <- temp_settl |> 
  group_by(indicator) |>
  filter(!!sym(sttl) >= 0.5) |>
  slice_max(`Priority.(opposite.of.severity)`)
#case 2 - none of options is 50%+
case2 <- temp_settl |> anti_join(case1, by="indicator")

av <- case2 |> 
  filter(!is.na(!!sym(sttl))) |>
  filter(!is.na(`Priority.(opposite.of.severity)`)) |>
  group_by(indicator) |>
  summarise(mean_p = round(mean(`Priority.(opposite.of.severity)`)))
# option with mean priority is taken
case2 <- case2 |>
  left_join(av) |>
  filter(`Priority.(opposite.of.severity)` == mean_p)

temp_aggr_ord <- plyr::rbind.fill(case1, case2)

temp_aggr_ord <- temp_aggr_ord |> select(indicator, var)

temp_aggr_ord_t <- temp_aggr_ord |>
  pivot_wider(everything(), names_from = "indicator", values_from = "var")

temp_aggr_ord_t <- temp_aggr_ord_t |>
  mutate(settlement = paste0(sttl))

ds_ord_aggr <- plyr::rbind.fill(ds_ord_aggr, temp_aggr_ord_t)
}


```
Select one nominal scales

if one is prevailing, take this, if not = "unknown"

```{r}
tabs_so_p_nom <- tabs_so_p |> filter(scale == "nominal")

ds_nom_aggr <- NULL

for(sttl in settlements){
  #sttl <- settlements[2]
  
  temp_settl <- tabs_so_p_nom |>
  select(indicator, var, !!sym(sttl))
# max. % option
case1 <- temp_settl |> 
  group_by(indicator) |>
  slice_max(!!sym(sttl))

nchoi <- case1 |>
  count()

case <- left_join(case1, nchoi)

case_1 <- case |>
  filter(n == 1)
# if more than one option in one question has max value - unknown
case_2 <- case |>
  anti_join(case_1) |>
  group_by(indicator) |>
  summarise(var = "unknown")

temp_nom <- rbind(case_1, case_2) |> select(indicator, var)

temp_aggr_nom_t <- temp_nom |>
  pivot_wider(everything(), names_from = "indicator", values_from = "var")

temp_aggr_nom_t <- temp_aggr_nom_t |>
  mutate(settlement = paste0(sttl))

ds_nom_aggr <- plyr::rbind.fill(ds_nom_aggr, temp_aggr_nom_t)
}
```

Medians for numerics

```{r}
numerics <- questions |> filter(type == "integer") |>
  select(name) |>
  unlist() |>
  as.vector()

ds_meds <- ds |> 
  group_by(a_5_settlement) |>
  summarise(across(all_of(numerics), median, na.rm = T))

# income variable, different currency, splitted by currency

med_inc <- ds |>
  group_by(a_5_settlement, b_41_a_income_currency) |>
  summarise(med_inc = median(b_41_average_income, na.rm = T)) |>
  pivot_wider(names_from = b_41_a_income_currency, values_from = med_inc, names_prefix = "median_income_")

ds_meds <- ds_meds |>select(-b_41_average_income) |>
  full_join(med_inc)
  
```
Joining data
```{r}
ds_aggr <- full_join(ds_ord_aggr, ds_nom_aggr) |>
  full_join(ds_sm_aggr) |>
  full_join(ds_meds, by = c("settlement" = "a_5_settlement"))
```
Recording the dataset
```{r}
write.xlsx(ds_aggr, "AoK_23_aggregated_upd_230913.xlsx")
write.xlsx(tabs, "settlement_tabs1.xlsx")
```

